---
title: "Orange Tree Dataset Analysis"
date: 2024-07-01
author: Carter Lembo
format: 
  html:
    toc: true
---

## Intro

The Orange dataset or Growth of Orange Trees contains information about the growth of a set of Orange trees. I want to explore this data and use linear regression to analyze how the age of the tree effects its circumference. 

## Setup

I will be using the tidyverse for data manipulation and visualization. 

```{r}
#| label: setup
#| output:  false
#| 
library(tidyverse)
```


## Exploration

Looking at the data we can see there are three columns. "Tree" identifies which tree is being measured, "age" measures the trees age in days, and "circumference" is the circumference of the tree in millimeters. 

```{r}
#| label: exploration
 
Orange |> 
  names()

Orange |> 
  head() |> 
  knitr::kable()

```

## Visualization

Before the linear regression analysis, I want to make sure the data is roughly linear.
```{r}
#| label: fig-orange-tree-growth
#| fig-cap: Orange Tree Growth
#| warning: false

Orange |> 
  ggplot(aes(x = age, y = circumference)) +
  geom_point(aes(color = Tree)) + 
  labs(
    title = "Orange Tree Growth with Tree Age",
    x = "Age (days)", y = "Circumference (mm)",
    caption = "Draper, N. R. and Smith, H. (1998), 
    Applied Regression Analysis (3rd ed), Wiley (exercise 24.N)."
  ) +
  scale_color_discrete(breaks = 1:5) +
  theme_bw()

```
 
@fig-orange-tree-growth shows that the data is approximately linear. 

## Issue

One issue with this data is that it is sampled many times from the same individuals, therefore the samples are not random or independent. This violates the assumptions the linear regression model so it would not be appropriate to use this model. Because of this, I will not be continuing using this data for now. 
