---
title: "Sleep Data Analysis"
date: 2024-06-27
author: Carter Lembo
format: html
---

## Intro

For this project, I want to analyse the built in dataset `datasets::sleep` or Student's Sleep Data. This dataset contains information about the increase in the number of hours of sleep of two different drugs.

## Setup

Using the tidyverse package for data wrangling and ggplot

```{r}
#| label: setup
#| warning: false
library(tidyverse)
```

## Exploration

Looking at the overall dataset where "extra" is the number of hours of sleep increase from a drug, "group" confusingly labels which drug is used as a treatment, and "ID" identifies which subject the treatment is applied to.

```{r}
#| label: exploration
sleep |> 
  names()

sleep |> 
  head() |> 
  knitr::kable()
```

Now I want to visualize the data for each drug.

```{r}
#| label: fig-sleep-distribution
#| fig-cap: Sleep Data Distributions
sleep |> 
  ggplot(aes(x = extra, y = group)) +
  geom_boxplot(aes(fill = group)) +
  geom_point() +
  labs(
    title = "Change in Hours of Sleep of Two Drugs",
    x = "Change in Hours Slept", y = "Drug Treatment",
    caption = "Source: Cushny, A. R. and Peebles, A. R. (1905) 
    The action of optical isomers: II hyoscines. 
    The Journal of Physiology 32, 501–510."
  ) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_y_discrete(labels = c("Drug 1", "Drug 2")) + 
  scale_x_continuous(breaks = scales::pretty_breaks(n = 8))
  
```

From this plot we can see that drug 2 seems a bit better than drug 1 but I want to test if there is a significant difference.

## Test Plan

I want to test if there is a significant difference between the true mean change in hours slept for an individual given drug 1 vs drug 2. Because the data is paired by the person given the drug, I can perform a paired t-test to see if there is a significant difference in the change in sleep between the two treatments. I plan to use a significance level of 0.05.

Hypotheses:

$H_0: \mu_{\text{diff}} = 0$  
$H_a: \mu_{\text{diff}} \neq 0$

where $\mu_{\text{diff}}$ is the difference between the true mean change in hours slept for a person given drug 1 minus drug 2.

## Check Normality

Before I can use a paired t-test, I must first ensure the distribution of the differences between the change in hours slept for an individual given drug 1 vs drug 2. First I will set up a vector of the differences for each person.

```{r}
#| label: differences
differences <- sleep |> 
  arrange(group, ID) |> 
  with(extra[group == 1] - extra[group == 2])

```

Next I will check to see if the data is approximately normal. To do this I will perform a quick Shapiro-Wilk test for normality with the the traditional significance level of 0.05.

```{r}
#| label: normality-test
differences |> 
  shapiro.test()
```

This gives a p-value of 0.03334 which is less than 0.05 we can say the distributions of differences is approximately normal. 

## Performing the Test

Now we can perform the paired t-test.

```{r}
#| label: test
differences |> 
  t.test(
    alternative = "two.sided",
    mu = 0,
    conf.level = 0.95
  )
```

Because the p-value of 0.002833 is less than 0.05 I can say there is significant evidence that there is a difference between the true mean change in hours slept for an individual given drug 1 vs drug 2, and reject the null hypothesis. 
