---
title: "Nile River Annual Flow"
date: 2024-07-03
author: Carter Lembo
format: 
  html:
    toc: true
---

## Intro

I want to estimate the average annual flow of the river Nile based on the data from 1871 to 1970 in the dataset Nile or Flow of the River Nile.

## Setup

I will be using the tidyverse for data manipulation and visualization.

```{r}
#| label: setup
#| output: false

library(tidyverse)
  
```

## Convert to Dataframe

This data is stored as a vector with an associated time series so I want to convert to a dataframe to do my analysis. Values for annual flow are measured in 10^8^m^3^.

```{r}
#| label: forming-dataframe


nile_df <- data.frame(flow = as.vector(Nile), year =  as.vector(time(Nile)))
```

## Visualization 

```{r}
#| label: fig-annual-nile-flow
#| fig-cap: Annual Nile River Flow 1871-1970s


nile_df |> 
  ggplot(aes(x = year, y = flow)) +
  geom_line() +
  labs(
    title = "Nile River Annual Flow 1871-1970",
     x = "Year", 
    y = expression(paste(plain("Water Flow "), 10^8*m^3)),
    caption = "Becker, R. A., Chambers, J. M. and Wilks, A. R. (1988) 
    The New S Language. Wadsworth & Brooks/Cole."
  ) +
  theme_bw()

```

## Get Sample

The data I have is for all years from 1870-1970. I want to estimate the true average for this range, but to do this, I am going to extract a sample of the total dataset. I am going to take a smaller sample of the overall data and then estimate based on that. Since I have the full data as well, I can compare to the true mean of the data. 

```{r}
#| label: extract-sample

set.seed(5) 

sample <- nile_df |> 
  sample_n(10)
```

## Interval

I want to estimate the true mean Nile flow in the years 1871-1970. I am going to use a confidence interval with a confidence level of 95%. To construct this interval, I will use the `t.test()` function and take the interval.  

```{r}
#| label: construct-interval

results <- sample$flow |> 
  t.test(conf.level = 0.95)


```

This gives me a mean estimate of `r results$estimate` and an interval of (`r results$conf.int`).

## Comparison

Now that I have an interval of (`r results$conf.int`), I can compare to the true mean of the population. 

```{r}
mean(nile_df$flow)
```
In this case, the interval captures the true mean river flow from 1871-1970. Changing the seed value in [Get Sample] will effect the exact numbers. 


